---
title: "DACSS 603 - Final Project Check-In #2"
author: "Daniel Seriy"
format: 
  html: 
    embed-resources: true
    self-contained-math: true
---

# Research Question

## Question 1

Does GDP have a positive effect on renewable energy generation/consumption?

## Question 2

Does Nuclear energy generation/consumption affect renewabole energy and GDP per capita’s effect?


# Hypothesis

## Hypothesis 1

Null: GDP per Capita has no effect on renewable generation/consumption
Alternative: As GDP per Capita increases, renewable generation/consumption increases

## Hypothesis 2

Null: Nuclear consumption/generation has no effect on renewable generation/consumption
Alternative: Nuclear consumption/generation has a positive effect on renewable consumption/generation


# Read in Packages

```{r, echo = T}

library(ggplot2)
library(stats)
library(dplyr)
library(stargazer)
library(effects)

```

# Read in Data

```{r, echo = T}

data <- read.csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-06/owid-energy.csv')

```

# Manipulate Data

```{r, echo = T}

data1 <- data %>%
  mutate(gdp_per_capita = gdp/population) %>%
  na.omit()

View(data1)

```

# Linear Models - Generation

## Model Creation

```{r, echo = T}

lm_generation_1 <- lm(renewables_electricity ~ gdp_per_capita, data = data1)
summary(lm_generation_1)

lm_generation_2 <- lm(renewables_electricity ~ gdp_per_capita + nuclear_electricity, data = data1)
summary(lm_generation_2)

lm_generation_3 <- lm(renewables_electricity ~ gdp_per_capita * nuclear_electricity, data = data1)
summary(lm_generation_3)

stargazer(lm_generation_1, lm_generation_2, lm_generation_3, type = "text")

```

## Visualizations - Generation

### Model 1

```{r, echo = T}

plot(x = data1$gdp_per_capita, y = data1$renewables_electricity,
     main = "Model #1 - Generation", sub = "Renewable Generation ~ GDP per Capita",
     xlab = "GDP per Capita (USD)", ylab = "Renewable Generation (TWH)")
abline(lm_generation_1, col = "red", lwd = 5)

```

### Model 2

```{r, echo = T}

plot(x = data1$gdp_per_capita + data1$nuclear_electricity, y = data1$renewables_electricity,
     main = "Model #2 - Generation", sub = "Renewable Generation ~ GDP per Capita + Nuclear Generation",
     xlab = "GDP per Capita (USD)", ylab = "Renewable Generation (TWH)")
abline(lm_generation_2, col = "red", lwd = 5)
     
```

### Model 3

```{r, echo = T}

plot_model_3_gen <- predictorEffect("gdp_per_capita", lm_generation_3)
as.data.frame(plot_model_3_gen)
plot(plot_model_3_gen, main = "Model #3 - Generation", sub = "Renewable Generation ~ GDP per Capita * Nuclear Generation",
     xlab = "GDP per Capita (USD)", ylab = "Renewable Generation (TWH)")

```

# Linear Models - Consumption

## Model Creation

```{r, echo = T}

lm_consumption_1 <- lm(renewables_consumption ~ gdp_per_capita, data = data1)
summary(lm_consumption_1)

lm_consumption_2 <- lm(renewables_consumption ~ gdp_per_capita + nuclear_consumption, data = data1)
summary(lm_consumption_2)

lm_consumption_3 <- lm(renewables_consumption ~ gdp_per_capita * nuclear_consumption, data = data1)
summary(lm_consumption_3)

stargazer(lm_consumption_1, lm_consumption_2, lm_consumption_3, type = "text")

```

## Visualizations - Consumption

### Model 1

```{r, echo = T}

plot(x = data1$gdp_per_capita, y = data1$renewables_consumption,
     main = "Model #1 - Consumption", sub = "Renewable Consumption ~ GDP per Capita",
     xlab = "GDP per Capita (USD)", ylab = "Renewable Consumption (TWH)")
abline(lm_consumption_1, col = "red", lwd = 5)

```

### Model 2

```{r, echo = T}

plot(x = data1$gdp_per_capita + data1$nuclear_consumption, y = data1$renewables_consumption,
     main = "Model #2 - Consumption", sub = "Renewable Consumption ~ GDP per Capita + Nuclear Generation",
     xlab = "GDP per Capita (USD)", ylab = "Renewable Generation (TWH)")
abline(lm_consumption_2, col = "red", lwd = 5)

```

### Model 3

```{r, echo = T}

plot_model_3_cons <- predictorEffect("gdp_per_capita", lm_consumption_3)
as.data.frame(plot_model_3_cons)
plot(plot_model_3_cons, main = "Model #3 - Consumption", sub = "Renewable Consumption ~ GDP per Capita * Nuclear Consumption",
     xlab = "GDP per Capita (USD)", ylab = "Renewable Consumption (TWH)")
```





```{r, echo = T}

```

```{r, echo = T}

```

```{r, echo = T}

```

```{r, echo = T}

```

```{r, echo = T}

```

```{r, echo = T}

```

# Discussion

## GDP per Capita

### Generation

As GDP per Capita increases, so does renewable energy generation (Coef = 0.008). As well, the model is statistically significant (p-value = 0.001).

### Consumption

Consumption shows similar results to the Generation results, but the relationship between GDP per Capita and Renewable consumption is higher, but with a lower statistical significance.

Coef = 0.027, p-value = 0.004


## GDP per Capita + Nuclear

### Generation

The addition of Nuclear energy maintains the existing relationship between GDP per Capita and renewable generation.

Coef = 0.004, p-value =0.258

Nuclear generation shows positive relationship with renewable generation

Coef = 0.258 , p-value = 0.067

### Consumption

presents similar results as generation, but a stronger relationship for both explanatory variables.

GDP Per Capita
Coef = 0.012, p-value = 0.005

Nuclear consumption
Coef = 0.356, p-value = 0.074 

## GDP per Capita * Nuclear

In both cases, as nuclear consumption/generation increases, there is a positive effect on the growth of GDP per Capita’s effect on renewable consumption/generation.

# Conclusion

We will reject the null hypothesis due to their being an effect by our explanatory variable for both Hypotheses.

## Hypothesis #1

Hypothesis is correct if only taking GDP per Capita into account in model.

## Hypothesis #2

Hypothesis is correct. Nuclear has positive relationship and effect on renewable energy.